---
title: "Parse GenBank Data"
author: "Leann M. Biancani"
output: pdf_document
---
#Download Data from GenBank:

NCBI nucleotide database: http://www.ncbi.nlm.nih.gov/nuccore

##Search terms:

#### Hyperiidea[ORGN] ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S or S18) or (28S) or (H3)) NOT Environmental samples[ORGN]

Total # of sequences: 337

```{r, eval=FALSE}
TAXA = "Hyperiidea.5genes"
```

#### Amphipoda[ORGN] ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S or S18) or (28S) or (H3)) NOT Environmental samples[ORGN]

Total # of sequences: 19617

```{r, eval=FALSE}
TAXA = "Amphipoda.5genes"
```

#### Isopoda[ORGN] ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S or S18) or (28S) or (H3)) NOT Environmental samples[ORGN]

Total # of sequences: 9125

```{r, eval=FALSE}
TAXA = "Isopoda.5genes"
```

#### (Amphipoda[ORGN] or Isopoda[ORGN]) ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S or S18) or (28S) or (H3)) NOT Environmental samples[ORGN]

Total # of sequences: 28742

```{r, eval=FALSE}
TAXA = "Amphipoda.Isopoda.5genes"
```

##Read Data into R

```{r eval=FALSE}
fasta_filename=paste(TAXA,"fasta",sep=".")
genbank_filename=paste(TAXA,"gb",sep=".")

#read data into R
fasta=readLines(fasta_filename)
genbank=readLines(genbank_filename)
```

##Parse Data

```{r, eval=FALSE, tidy=TRUE}
#load stringr library
library(stringr)

#create vector of genbank data
genbank_vector=c()
gb=""
for(line in genbank){
  gb=paste(gb,line,sep=" ")
  if(line=="//"){
  #if(str_detect(line,"//")){
    gb=str_replace_all(gb," +"," ")
    genbank_vector=append(genbank_vector,gb)
    gb=""
  }
}

genelookup= data.frame(
  rawlist=c("[Cc]ytochrome [Cc ]{0,2}[Oo]xidase [subunit ]{0,8}[Ii1][^Ii1]","[Cc][Oo][Xx]{0,1}[Ii1][^Ii1]","16S","large subunit.*mitochondrial","small subunit","18S","S18","28S","[Hh]3","[Hh]istone 3","genome"),
  genelist=c("mt_COI","mt_COI","mt_16S","mt_16S","n_18S","n_18S","n_18S","n_28S","n_H3","n_H3","mt_genome")
)

#initiate vectors
fasta_header=c()
sequence=c()

species_name=c()
organism=c()
accession=c()
accession_gb=c()
description=c()

voucher=c()
country=c()
isolate=c()
gene=c()
bp_length=c()
taxon_id=c()

author_ref1=c()
title_ref1=c()
journal_ref1=c()

author_ref2=c()
title_ref2=c()
journal_ref2=c()

author_ref3=c()
title_ref3=c()
journal_ref3=c()

others=c()

#parse fasta (& genbank) data
seq=""
record=0
for (i in seq(length(fasta))) {
  line = fasta[i]
  # if a header is detected:
  if (str_detect(line,fixed(">"))){
    #if sequence was recorded for previous record, add it to the sequence vector
    if (i>1) {sequence=append(sequence,seq)}
    seq="" #reset seq
    
    record=record+1 #track record number
    
    #store fasta header
    fasta_header = append(fasta_header,line)
    
    #parse fasta header
    header_split=strsplit(line, "|", fixed=TRUE)
    
    #accession from fasta
    accession=append(accession,header_split[[1]][4])
    
    #description from fasta
    des=str_trim(header_split[[1]][5], side="left")
    description=append(description,des)
    
    #species name from fasta
    #[cfa. ]{0,5} added to accomodate "Genus cf. species" and "Genus aff. species"
    species_name=append(species_name,str_extract(des,"[A-Z][a-z]*[ .][cfa. ]{0,5}[a-z.]+"))
    
    #gene from fasta
    ingenes=FALSE
    gen=""
    for(i in seq(length(genelookup$rawlist))){
      if(str_detect(des,as.character(genelookup$rawlist[i]))){
        lookup=as.character(genelookup$genelist[i])
        if(!str_detect(gen,lookup)){gen=paste(lookup,gen,sep=" ")}
        ingenes=TRUE
      }
    }
    if(!ingenes){
      gen = "other"
      others=append(others,des)
    }
    gene=append(gene,gen)
    
    
    
    #parse data from genbank file:
    gb=genbank_vector[record]
    
    #accession from gb
    accession_pattern="ACCESSION ([A-Z]{2}\\d{6})"
    acc_gb=str_replace(str_extract(gb,accession_pattern),accession_pattern,"\\1")
    accession_gb=append(accession_gb,acc_gb)
    
    #organism from gb
    org_pattern="organism=\"([^\"]+)\""
    org=str_replace(str_extract(gb,org_pattern),org_pattern,"\\1")
    organism=append(organism,org)
    
    #voucher number from gb
    vou_pattern="voucher=\"([^\",]*\\d[^\",]*)\""
    vou=str_replace(str_extract(gb,vou_pattern),vou_pattern,"\\1")
    voucher=append(voucher,vou)
    
    #country from gb
    country_pat="country=\"([^\"]+)\""
    coun=str_replace(str_extract(gb,country_pat),country_pat,"\\1")
    country=append(country,coun)
    
    #isolate number from gb
    iso_pattern="/isolate=\"([^\"]+)\""
    iso=str_replace(str_extract(gb,iso_pattern),iso_pattern,"\\1")
    isolate=append(isolate,iso)
    
    #sequence length from gb
    bps=as.numeric(str_extract(str_extract(gb,"[:digit:]+ bp"),"[:digit:]+"))
    bp_length=append(bp_length,bps)
    
    #taxon ID from gb
    tax=as.numeric(str_extract(str_extract(gb,"taxon:[:digit:]+"),"[:digit:]+"))
    taxon_id=append(taxon_id,tax)
    
    #author and/or consortium for References 1-3 from gb
    reference_ends=c("REFERENCE","COMMENT","FEATURES")
    for(end in reference_ends){gb=str_replace_all(gb,end,paste("@",end,sep=""))}
    reference_keys=c("AUTHORS","CONSRTM","TITLE","JOURNAL")
    for(key in reference_keys){gb=str_replace_all(gb,key,paste("%",key,sep=""))}
    
    #REFERENCES 1-3
    for (i in 1:3) {
      auth_pat=paste("REFERENCE ","[^@]+%AUTHORS ([^%]+) %",sep=as.character(i))
      cons_pat=paste("REFERENCE ","[^@]+%CONSRTM ([^%]+) %",sep=as.character(i))
      title_pat=paste("REFERENCE ","[^@]+%TITLE ([^%]+) %",sep=as.character(i))
      journ_pat=paste("REFERENCE ","[^@]+%JOURNAL ([^%@]+) [%@]",sep=as.character(i))
      auth= str_replace(str_extract(gb,auth_pat),auth_pat,"\\1")
      cons= str_replace(str_extract(gb,cons_pat),cons_pat,"\\1")
      title= str_replace(str_extract(gb,title_pat),title_pat,"\\1")
      journ= str_replace(str_extract(gb,journ_pat),journ_pat,"\\1")
      if(is.na(auth)){
        auth=cons # in case consortium listed but no author
      }else if(!is.na(cons)){
        auth=paste(auth,cons,sep=", ") # in case author & consortium listed (never?)
      }
      if(i==1){
        author_ref1=append(author_ref1,auth)
        title_ref1=append(title_ref1,title)
        journal_ref1=append(journal_ref1,journ)
      }else if(i==2){
        author_ref2=append(author_ref2,auth)
        title_ref2=append(title_ref2,title)
        journal_ref2=append(journal_ref2,journ)
      }else if(i==3){
        author_ref3=append(author_ref3,auth)
        title_ref3=append(title_ref3,title)
        journal_ref3=append(journal_ref3,journ)
      }
    }
    
  } else {
    # if the line is not a header, add the line to seq
    seq=paste(seq,line,sep="")
  }
}
sequence=append(sequence,seq)

df=data.frame(species_name,organism,taxon_id,voucher,isolate,country,accession,accession_gb,gene,bp_length,author_ref1,title_ref1,journal_ref1,author_ref2,title_ref2,journal_ref2,author_ref3,title_ref3,journal_ref3,description,fasta_header,sequence,stringsAsFactors=FALSE)

print(others)

```

##Build Gene Table

```{r, eval=FALSE, tidy=TRUE}

genelist=append((unique(as.character(genelookup$genelist))),"other")

#initialize gene vectors:
for(gene in genelist) {assign(gene,c())}

g=""
for(i in seq(nrow(df))){
  entry=as.character(df$gene[i])
  for(gene in genelist){
    if(str_detect(entry,gene)){g=df$accession[i]}
    assign(gene,do.call(append,list(as.name(gene),g)))
    g=""
  }
}

gene_table=df[,c(1:5,11)]
gene_table=data.frame(gene_table,do.call(cbind,lapply(genelist,as.name)),stringsAsFactors = FALSE)

#aggregate rows where voucher # matches (omits rows where voucher is NA)
#gt=aggregate(gene_table, by=list(gene_table$voucher), max, na.rm=TRUE)
gt=aggregate(gene_table, by=list(gene_table$voucher), FUN=unique)

#remove Group.1 row
gt=gt[,c(2:ncol(gt))]

#replace rows where voucher is NA
gt=rbind(gt,gene_table[is.na(voucher),])
```

##Build TaxonID table

```{r, eval=FALSE}

ID_table=gene_table[,c(1:3,7:ncol(gene_table))]

ID=aggregate(ID_table, by=list(ID_table$taxon_id), unique)

#remove Group.1 row
ID=ID[,c(2:ncol(ID))]

```

##Output Data

```{r, eval=FALSE, tidy=TRUE}
#parsed data table
df_filename=paste(TAXA,DATABASE,"alldata.txt",sep=".")
write.table(df, df_filename, sep="\t",row.names=FALSE)
#gene table
gt_filename=paste(TAXA,DATABASE,"genetable.txt",sep=".")
write.table(gt, gt_filename, sep="\t",row.names=FALSE)

```
