---
title: "Parse GenBank Data"
author: "Leann M. Biancani"
output: pdf_document
---
#Download Data from GenBank:

NCBI nucleotide database: http://www.ncbi.nlm.nih.gov/nuccore

##Search terms:

#### Hyperiidea[ORGN] ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S or S18) or (28S) or (H3)) NOT Environmental samples[ORGN]

Total # of sequences: 337

```{r, eval=FALSE}
#TAXA = "Hyperiidea.5genes"
```

#### Amphipoda[ORGN] ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S or S18) or (28S) or (H3)) NOT Environmental samples[ORGN]

Total # of sequences: 19617

```{r, eval=FALSE}
TAXA = "Amphipoda.5genes"
```

#### Isopoda[ORGN] ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S or S18) or (28S) or (H3)) NOT Environmental samples[ORGN]

Total # of sequences: 9125

```{r, eval=FALSE}
#TAXA = "Isopoda.5genes"
```

#### (Amphipoda[ORGN] or Isopoda[ORGN]) ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S or S18) or (28S) or (H3)) NOT Environmental samples[ORGN]

Total # of sequences: 28742 (29858 on 9/2/16)

10/28/16: filter out "PREDICTED" (57 bad sequences "predicted" to be H. azteca out of 29956 total sequences) = 29899 sequences

(Amphipoda[ORGN] or Isopoda[ORGN]) ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S or S18) or (28S) or (H3)) NOT (Environmental samples[ORGN] or PREDICTED)

10/28/16: filter out genome[Title] because we're not looking at whole mt genomes (113 genome sequences out of 29899 total sequences) = 29786

(Amphipoda[ORGN] or Isopoda[ORGN]) ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S or S18) or (28S) or (H3)) NOT (Environmental samples[ORGN] or PREDICTED or genome[Title])

10/20/16: filter out UNVERIFIED sequences (103 sequences out of 29786 total) 
= 29683 sequences downloaded (sorted by accession, include GI # (impacts format of fasta header!))

11/4/2016 - filter out 454 data (keep coming across poor sequences that report being sequenced with 454) - filters 271 of 29944 sequences available on 11/4
= 29673 sequences

#### (Amphipoda[ORGN] or Isopoda[ORGN]) ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S or S18) or (28S) or (H3)) NOT (Environmental samples[ORGN] or PREDICTED or genome[Title] or UNVERIFIED or 454)

```{r, eval=FALSE}
TAXA = "Amphipoda.Isopoda.5genes"
```

11/4/2016 - Split Amphipoda and Isopoda search and parse in order to filter Isopoda for selected taxa:
(New DataParse_Isopods.Rmd created to parse Isopod data and create fasta file for each gene - will be concatinated with Amphipoda fasta files later.)
11/4/2016 - 20179 sequences downloaded (sorted by accession, include GI # (impacts format of fasta header!))

11/5/2016 - Filter to remove "internal transcribed spacer" (filters 655 sequences = 19524 sequences)
These sequences are long and contain very small pieces of multiple genes and are not aligning well.
11/5/2016 - 19524 sequences downloaded (sorted by accession, include GI # (impacts format of fasta header!))

#### Amphipoda[ORGN] ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S or S18) or (28S) or (H3)) NOT (Environmental samples[ORGN] or PREDICTED or genome[Title] or UNVERIFIED or 454 or "internal transcribed spacer")

1/26/17
remove Unclassified Amphipoda: txid555092[Organism:exp] (696 total sequences)
#### Amphipoda[ORGN] ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S or S18) or (28S) or (H3)) NOT (Environmental samples[ORGN] or PREDICTED or genome[Title] or UNVERIFIED or 454 or "internal transcribed spacer") NOT txid555092[Organism:exp]
20104 sequences on 1/26/17

remove "unclassified" (removes 43 Chiltoniidae sp.)
#### Amphipoda[ORGN] ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S or S18) or (28S) or (H3)) NOT (Environmental samples[ORGN] or PREDICTED or genome[Title] or UNVERIFIED or 454 or "internal transcribed spacer" or txid555092[Organism:exp] or unclassified) 
20061 sequences on 1/26/17
= 20061 sequences downloaded (sorted by accession, include GI # (impacts format of fasta header!))
spreadsheet of data created for this dataset

S18 does not equal 18S (Removes 1 sequence)
#### Amphipoda[ORGN] ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S) or 28S or (H3)) NOT (Environmental samples[ORGN] or PREDICTED or genome[Title] or UNVERIFIED or 454 or "internal transcribed spacer" or txid555092[Organism:exp] or unclassified)
20244 sequences on 2/13/17
= 20244 sequences downloaded (sorted by accession, include GI # (impacts format of fasta header!))

#### Amphipoda[ORGN] ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S) or 28S or (H3)) NOT (Environmental samples[ORGN] or PREDICTED or genome[Title] or UNVERIFIED or 454 or "internal transcribed spacer" or txid555092[Organism:exp] or unclassified)
Same as above but downloaded on 3/15/2017
(Added H3 and linked COI for Platorchestia platensis to selected list as well)
= 20608 sequences downloaded (sorted by accession, include GI # (impacts format of fasta header!))

#### Amphipoda[ORGN] ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S) or 28S or (H3)) NOT (Environmental samples[ORGN] or PREDICTED or genome[Title] or UNVERIFIED or "internal transcribed spacer" or txid555092[Organism:exp] or unclassified)
Replace 454 sequences (262 sequences) some also added to selected list
add new sequences 3/15/2017-4/17/2017 (796 sequences addes) some also added to selected list
= 21743 sequences downloaded on 4/17/2017 (sorted by accession, include GI # (impacts format of fasta header!))
add new sequences 4/17/2017-5/5/2017 - some added to selected list as well
= 21765 sequences downloaded on 5/5/2017 (sorted by accesssion, include GI # (impacts format of fasta header!))
5/11/17 none to add
7/25/17 = 22822 (some added to selected list)
7/26/17 = 22822

stop excluding UNVERIFIED sequences:
#### Amphipoda[ORGN] ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S) or 28S or (H3)) NOT (Environmental samples[ORGN] or PREDICTED or genome[Title] or "internal transcribed spacer" or txid555092[Organism:exp] or unclassified)
add 5 sequences, 2 added to selected list
7/26/17 - 22827


```{r eval=FALSE}
TAXA = "Amphipoda.5genes"
```

#### Isopoda[ORGN]  ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S) or 28S or (H3)) NOT (Environmental samples[ORGN] or PREDICTED or genome[Title] or UNVERIFIED or 454 or "internal transcribed spacer" or txid555092[Organism:exp] or unclassified)
9929 sequences on 3/19/17
= 9929 sequences downloaded (sorted by accession, include GI # (impacts format of fasta header!))
= 11524 sequences on 7/24/17

#### Isopoda[ORGN]  ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S) or 28S or (H3)) NOT (Environmental samples[ORGN] or PREDICTED or genome[Title] or UNVERIFIED or "internal transcribed spacer" or txid555092[Organism:exp] or unclassified)
Including 454 sequences
= 11533 sequences on 7/24/17
= 11533 sequences on 7/26/17

#### Isopoda[ORGN]  ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S) or 28S or (H3)) NOT (Environmental samples[ORGN] or PREDICTED or genome[Title] or UNVERIFIED or "internal transcribed spacer" or txid555092[Organism:exp] or unclassified)

Stop excluding UNVERIFIED sequences:
#### Isopoda[ORGN] ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S) or 28S or (H3)) NOT (Environmental samples[ORGN] or PREDICTED or genome[Title] or "internal transcribed spacer" or txid555092[Organism:exp] or unclassified)
adds ~100 sequences, a couple of H3 seqs added to selected list
= 11638 on 7/26/17

```{r eval=FALSE}
TAXA = "Isopoda.5genes"
```


#### (Amphipoda[ORGN] or Isopoda[ORGN]) ((coi or co1 or coxi or cox1 or "cytochrome oxidase subunit 1" or "cytochrome oxidase subunit i" or "cytochrome oxidase i" or "cytochrome c oxidase subunit 1" or "cytochrome c oxidase subunit i") or (("large subunit" mitochondrial) or 16S) or (("small subunit" not mitochondrial) or 18S) or 28S or (H3)) NOT (Environmental samples[ORGN] or PREDICTED or genome[Title] or "internal transcribed spacer" or txid555092[Organism:exp] or unclassified)
8/8/17 = 34555 sequences downloaded (none added to selected list)

```{r eval=FALSE}
TAXA = "Amphipoda.Isopoda.5genes"
```

##Read Data into R

```{r eval=FALSE}
fasta_filename=paste(TAXA,"fasta",sep=".")
genbank_filename=paste(TAXA,"gb",sep=".")

#read data into R
fasta=readLines(fasta_filename)
genbank=readLines(genbank_filename)
```

##Parse Data

```{r, eval=FALSE, tidy=TRUE}
#load stringr library
library(stringr)

#create vector of genbank data
genbank_vector=c()
gb=""
for(line in genbank){
  gb=paste(gb,line,sep=" ")
  if(line=="//"){
  #if(str_detect(line,"//")){
    gb=str_replace_all(gb," +"," ")
    genbank_vector=append(genbank_vector,gb)
    gb=""
  }
}

# genelookup= data.frame(
#   rawlist=c("[Cc]ytochrome [Cc ]{0,2}[Oo]xidase [subunit ]{0,8}[Ii1][^Ii1]","[Cc][Oo][Xx]{0,1}[Ii1][^Ii1]","16S","large subunit.*mitochondrial","small subunit","18S","S18","28S","[Hh]3","[Hh]istone 3","mito.*genome"),
#   genelist=c("mt_COI","mt_COI","mt_16S","mt_16S","n_18S","n_18S","n_18S","n_28S","n_H3","n_H3","mt_genome")
# )

genelookup= data.frame(
  rawlist=c("[Cc]ytochrome [Cc ]{0,2}[Oo]xidase [subunit ]{0,8}[Ii1][^Ii1]","[Cc][Oo][Xx]{0,1}[Ii1][^Ii1]","16S","18S","28S","[Hh]3","[Hh]istone 3"),
  genelist=c("mt_COI","mt_COI","mt_16S","n_18S","n_28S","n_H3","n_H3")
)
# S18 does not equal 18S

#large/small subunit designations have been ambiguous - will look for explicit 16/18/28S designations

#initiate vectors
fasta_header=c()
sequence=c()

species_name=c()
organism=c()
accession=c()
accession_gb=c()
description=c()

voucher=c()
country=c()
isolate=c()
gene=c()
bp_length=c()
taxon_id=c()

author_ref1=c()
title_ref1=c()
journal_ref1=c()

author_ref2=c()
title_ref2=c()
journal_ref2=c()

author_ref3=c()
title_ref3=c()
journal_ref3=c()

others=c()

#parse fasta (& genbank) data
dna=""
record=0
for (i in seq(length(fasta))) {
  line = fasta[i]
  # if a header is detected:
  if (str_detect(line,fixed(">"))){
    #if sequence was recorded for previous record, add it to the sequence vector
    if (i>1) {sequence=append(sequence,dna)}
    dna="" #reset dna

    record=record+1 #track record number

    #store fasta header
    fasta_header = append(fasta_header,line)

    #parse fasta header
    header_split=strsplit(line, "|", fixed=TRUE)

    #accession from fasta
    accession=append(accession,header_split[[1]][4])

    #description from fasta
    des=str_trim(header_split[[1]][5], side="left")
    description=append(description,des)

    #species name from fasta
    #[cfa. ]{0,5} added to accomodate "Genus cf. species" and "Genus aff. species"
    species_name=append(species_name,str_extract(des,"[A-Z][a-z]*[ .][cfa. ]{0,5}[a-z.]+"))

    #parse data from genbank file:
    gb=genbank_vector[record]    
    
    #gene from genbank
    # was pulling gene name from fasta description but other elements of description were being flagged as gene names (ex. H3)
    
    gen_pattern="gene=\"[^\"]+\""
    #extract all listed genes
    rawgen=paste(str_extract_all(gb,gen_pattern)[[1]],collapse="")
    pro_pattern="product=\"[^\"]+\""
    #extract all listed products
    rawpro=paste(str_extract_all(gb,pro_pattern)[[1]],collapse="")
    rawgenpro=paste(rawgen,rawpro,collapse = "")
    
    #check rawgenpro for vague "subunit" designations
    if(str_detect(gb,"mitochondrial")){
      if(str_detect(rawgenpro,"large subunit")){rawgenpro=paste(rawgenpro,"16S",sep=" ")}
      #if(str_detect(rawgenpro,"small subunit")){} #not needed as all are labeled 16S
    }else{
      if(str_detect(rawgenpro,"small subunit")){rawgenpro=paste(rawgenpro,"18S",sep="")}
      if(str_detect(rawgenpro,"large subunit")){} #not needed as all are labeled 28S
    }

    #if no info in "gene" or "product", check for "note":
    if(str_length(rawgenpro)<2){
      note_pattern="note=\"[^\"]+\""
      rawnote=paste(str_extract_all(gb,note_pattern)[[1]],collapse="")
      rawgenpro=rawnote
    }

    ingenes=FALSE
    gen=""
    for(i in seq(length(genelookup$rawlist))){
      if(str_detect(rawgenpro,as.character(genelookup$rawlist[i]))){
        lookup=as.character(genelookup$genelist[i])
        if(!str_detect(gen,lookup)){gen=paste(lookup,gen,sep=" ")}
        ingenes=TRUE
      }
    }
    if(!ingenes){
      gen = "other"
      others=append(others,des)
    }
    gene=append(gene,gen)





    #accession from gb
    accession_pattern="ACCESSION ([A-Z]{1,2}\\d{5,6})"
    acc_gb=str_replace(str_extract(gb,accession_pattern),accession_pattern,"\\1")
    accession_gb=append(accession_gb,acc_gb)

    #organism from gb
    org_pattern="organism=\"([^\"]+)\""
    org=str_replace(str_extract(gb,org_pattern),org_pattern,"\\1")
    organism=append(organism,org)

    #voucher number from gb
    vou_pattern="voucher=\"([^\",]*\\d[^\",]*)\""
    vou=str_replace(str_extract(gb,vou_pattern),vou_pattern,"\\1")
    if(!is.na(vou)){
      if(str_detect(vou,"collection")){vou=NA}
    }
    voucher=append(voucher,vou)

    #country from gb
    country_pat="country=\"([^\"]+)\""
    coun=str_replace(str_extract(gb,country_pat),country_pat,"\\1")
    country=append(country,coun)

    #isolate number from gb
    iso_pattern="/isolate=\"([^\"]+)\""
    iso=str_replace(str_extract(gb,iso_pattern),iso_pattern,"\\1")
    isolate=append(isolate,iso)

    #sequence length from gb
    bps=as.numeric(str_extract(str_extract(gb,"[:digit:]+ bp"),"[:digit:]+"))
    bp_length=append(bp_length,bps)

    #taxon ID from gb
    tax=as.numeric(str_extract(str_extract(gb,"taxon:[:digit:]+"),"[:digit:]+"))
    taxon_id=append(taxon_id,tax)

    #author and/or consortium for References 1-3 from gb
    reference_ends=c("REFERENCE","COMMENT","FEATURES")
    for(end in reference_ends){gb=str_replace_all(gb,end,paste("@",end,sep=""))}
    reference_keys=c("AUTHORS","CONSRTM","TITLE","JOURNAL")
    for(key in reference_keys){gb=str_replace_all(gb,key,paste("%",key,sep=""))}

    #REFERENCES 1-3
    for (i in 1:3) {
      auth_pat=paste("REFERENCE ","[^@]+%AUTHORS ([^%]+) %",sep=as.character(i))
      cons_pat=paste("REFERENCE ","[^@]+%CONSRTM ([^%]+) %",sep=as.character(i))
      title_pat=paste("REFERENCE ","[^@]+%TITLE ([^%]+) %",sep=as.character(i))
      journ_pat=paste("REFERENCE ","[^@]+%JOURNAL ([^%@]+) [%@]",sep=as.character(i))
      auth= str_replace(str_extract(gb,auth_pat),auth_pat,"\\1")
      cons= str_replace(str_extract(gb,cons_pat),cons_pat,"\\1")
      title= str_replace(str_extract(gb,title_pat),title_pat,"\\1")
      journ= str_replace(str_extract(gb,journ_pat),journ_pat,"\\1")
      if(is.na(auth)){
        auth=cons # in case consortium listed but no author
      }else if(!is.na(cons)){
        auth=paste(auth,cons,sep=", ") # in case author & consortium listed (never?)
      }
      if(i==1){
        author_ref1=append(author_ref1,auth)
        title_ref1=append(title_ref1,title)
        journal_ref1=append(journal_ref1,journ)
      }else if(i==2){
        author_ref2=append(author_ref2,auth)
        title_ref2=append(title_ref2,title)
        journal_ref2=append(journal_ref2,journ)
      }else if(i==3){
        author_ref3=append(author_ref3,auth)
        title_ref3=append(title_ref3,title)
        journal_ref3=append(journal_ref3,journ)
      }
    }

  } else {
    # if the line is not a header, add the line to dna
    dna=paste(dna,line,sep="")
  }
}
sequence=append(sequence,dna)

df=data.frame(species_name,organism,taxon_id,voucher,isolate,country,accession,accession_gb,gene,bp_length,author_ref1,title_ref1,journal_ref1,author_ref2,title_ref2,journal_ref2,author_ref3,title_ref3,journal_ref3,description,fasta_header,sequence,stringsAsFactors=FALSE)

head(others)
print(length(others))
df_others<-df[gene=="other",]
```

##Build Gene Table

```{r, eval=FALSE, tidy=TRUE}

genelist=append((unique(as.character(genelookup$genelist))),"other")

#initialize gene vectors:
for(gene in genelist) {assign(gene,c())}

g=""
for(i in seq(nrow(df))){
  entry=as.character(df$gene[i])
  for(gene in genelist){
    if(str_detect(entry,gene)){g=df$accession[i]}
    assign(gene,do.call(append,list(as.name(gene),g)))
    g=""
  }
}
#added 10 to line below to add bp_lentgh to gene_table
gene_table=df[,c(1:5,10,11)]
gene_table=data.frame(gene_table,do.call(cbind,lapply(genelist,as.name)),stringsAsFactors = FALSE)
```

##Build Fasta File for Each Gene

```{r, eval=FALSE}
#define function create.fasta
create.fasta <- function(dataname,dataframe) {
  filename=paste(dataname,"fasta",sep=".")
  fasta_file=file(filename,"at")
  locateColNames=function(x){which(colnames(dataframe)==x)}
  headindex=sapply(c("taxon_id","organism","accession"),locateColNames)
  seqindex=locateColNames("sequence")
  for (i in seq(nrow(dataframe))){
    if(i==1){begin=">"}else{begin="\n>"}
    gene_fasta=c(begin,sapply(as.character(dataframe[i,headindex]),paste,"|",sep=""),"\n",dataframe[i,seqindex])
    #remove spaces and . from names
    gene_fasta[3]=str_replace_all(gene_fasta[3]," ","_")
    gene_fasta[3]=str_replace_all(gene_fasta[3],fixed("."),"")
    writeLines(gene_fasta,con=fasta_file,sep="")
  }
  close(fasta_file)
}


for (gene in genelist) {
  dataname=paste(TAXA,gene,sep=".")
  #extract rows from df where accession number is in gene vector:
  gene_df=df[as.logical(do.call(match,list(df$accession,as.name(gene),0))),]
  create.fasta(dataname,gene_df)
}

``` 

##Create Hurt & Hurt list
```{r, eval=FALSE}
HurtHouList=readLines("hurt_hou_accessions.txt")
HurtHouList=str_split(HurtHouList," ")
HurtHouList=unique(HurtHouList[[1]])
```

##Build Fasta File for Hou & Hurt Data

```{r, eval=FALSE}
## for testing, backup df:
#df_backup=df
## create smaller df:
#df=df[seq(1,nrow(df),100),]

for (gene in genelist) {
  #extract rows from df where genbank accession number is in gene vector:
  gene_df=df[as.logical(do.call(match,list(df$accession,as.name(gene),0))),]
  #extract rows from gene_df where genbank accession number is in HurtHouList vector:
  HH_gene_df=gene_df[gene_df$accession_gb %in% HurtHouList,]
  filename=paste("HH_",gene,".fasta",sep="")
  fasta_file=file(filename,"at")
  
  locateColNames=function(x){which(colnames(HH_gene_df) == x)}
  #headindex=sapply(c("species_name","accession_gb"),locateColNames)
  nameindex=locateColNames("species_name")
  accessionindex=locateColNames("accession_gb")
  seqindex=locateColNames("sequence")
  
  for (i in seq(nrow(HH_gene_df))){
    if(i==1){begin=">"}else{begin="\n>"}
    #gene_fasta=c(begin,sapply(as.character(HH_gene_df[i,headindex]),paste,"|",sep=""),"\n",HH_gene_df[i,seqindex])
    gene_fasta=c(begin,HH_gene_df[i,nameindex],"|",HH_gene_df[i,accessionindex],"\n",HH_gene_df[i,seqindex])
    #remove spaces and . from names
    gene_fasta[2]=str_replace_all(gene_fasta[2]," ","_")
    gene_fasta[2]=str_replace_all(gene_fasta[2],fixed("."),"")
  writeLines(gene_fasta,con=fasta_file,sep="")
  }
  
  close(fasta_file)
}

```

##Create SICB list
```{r, eval=FALSE}
sicbList=readLines("SICB_accessions.txt")
sicbList=str_split(sicbList,",")
sicbList=unique(sicbList[[1]])
```

##Build Fasta File for SICB list

```{r, eval=FALSE}
## for testing, backup df:
#df_backup=df
## create smaller df:
#df=df[seq(1,nrow(df),100),]

for (gene in genelist) {
  #extract rows from df where genbank accession number is in gene vector:
  gene_df=df[as.logical(do.call(match,list(df$accession,as.name(gene),0))),]
  #extract rows from gene_df where genbank accession number is in sicbList vector:
  sicb_gene_df=gene_df[gene_df$accession_gb %in% sicbList,]
  filename=paste("sicb_",gene,".fasta",sep="")
  fasta_file=file(filename,"at")
  
  locateColNames=function(x){which(colnames(sicb_gene_df) == x)}
  #headindex=sapply(c("species_name","accession_gb"),locateColNames)
  nameindex=locateColNames("species_name")
  accessionindex=locateColNames("accession_gb")
  seqindex=locateColNames("sequence")
  
  for (i in seq(nrow(sicb_gene_df))){
    if(i==1){begin=">"}else{begin="\n>"}
    gene_fasta=c(begin,sicb_gene_df[i,nameindex],"|",sicb_gene_df[i,accessionindex],"\n",sicb_gene_df[i,seqindex])
    #remove spaces and . from names
    gene_fasta[2]=str_replace_all(gene_fasta[2]," ","_")
    gene_fasta[2]=str_replace_all(gene_fasta[2],fixed("."),"")
  writeLines(gene_fasta,con=fasta_file,sep="")
  }
  
  close(fasta_file)
}

```

##Build TaxonID table

```{r, eval=FALSE}

ID_table=gene_table[,c(1:3,7:ncol(gene_table))]

ID=aggregate(ID_table, by=list(ID_table$taxon_id), unique)

#remove Group.1 row
ID=ID[,c(2:ncol(ID))]

#turn lists into character strings & cleanup:
for(i in seq(nrow(ID))){
  for(j in seq(ncol(ID))){
    cell = ID[i,j]
    if(!is.na(cell)){
      cell = paste(c(cell[[1]]), collapse=", ")
      cell = str_replace(cell,", $","")
      cell = str_replace(cell,"^, ","")
      ID[i,j]=(cell)
    }
  }
}

ID[] = lapply(ID, as.character)
ID[] = transform(ID, taxon_id = as.numeric(taxon_id))
```

##Output Data

```{r, eval=FALSE, tidy=TRUE}
#parsed data table
df_filename=paste(TAXA,"alldata.txt",sep=".")
write.table(df, df_filename, sep="\t",row.names=FALSE)
#gene table
gt_filename=paste(TAXA,"genetable.txt",sep=".")
write.table(gene_table, gt_filename, sep="\t",row.names=FALSE)
#taxonID table
id_filename=paste(TAXA,"taxonIDtable.txt",sep=".")
write.table(ID, id_filename, sep="\t",row.names=FALSE)
```

##Taxon Search


```{r, eval=FALSE, tidy=TRUE}
library(stringr)
#Generate a list of Genera
organism_list=unique(organism)
genus_pattern="([^ ]*)"
genera=c()
for(i in organism_list){
  genus=str_extract(i, genus_pattern)
  genera=append(genera,genus)
}
genera=unique(genera)
```

```{r, eval=FALSE, tidy=TRUE}
test=c(
"Califarcturus","Spinarcturus","Thermoarcturus"
)
count=0
for(i in test){
  if(i %in% genera){
    print(i)
    count=count+1
  }
}
print(count)
```


```{r, eval=FALSE, tidy=TRUE}

my_genera=readLines("mygenera.txt")

count=0
for(i in genera){
  if(!(i %in% my_genera)){
    print(i)
    count=count+1
  }
}

print(count)
```


```{r, eval=FALSE, tidy=TRUE}

df_COI=df[!(is.na(str_match(df$gene,"mt_COI"))),]#######
lenCOI=df_COI$bp_length
densityCOI=density(lenCOI)
plot(densityCOI)

df_28S=df[!(is.na(str_match(df$gene,"n_28S"))),]
len28S=df_28S$bp_length
density28S=density(len28S)
plot(density28S)

df_28S_long1200=subset(df_28S,bp_length>=1200)
df_28S_frag1200=subset(df_28S,bp_length<1200)



```